Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-31 introduction.tex
proof ch 1

However, \emph{seroprevalence} surveys are more feasible.
Seroprevalence is the proportion of individuals in the population which have antibodies against a particular infection, indicating that the individual was previously infected~\autocite{cdcSeroprevalence}.
Many more individuals have ever been infected than are currently infected, therefore, seroprevalence surveys can be powered much more cheaply~\autocite{wuSeroprevSimulation}.
However, seroprevalence is a delayed measure because there is a lag between an individual being infected and them developing antibodies.

@@

Next, are semi-mechanistic models.
These relate incidence in the past to current (and future) incidence but do not explicitly represent the susceptible population.
A common type of semi-mechanistic models is that based on a \emph{renewal equation}~\autocite{bhattSemimechanistic,fraserRenewal}.
In discrete time, it takes the form:
\begin{align}
    \E(Z_t \mid Z_0, \dots, Z_{t-1}) &= \sum_{\tau=0}^t \beta(t,\tau) Z_{t-\tau} d \tau
    \label{intro:eq:renewal}
\end{align}
where $Z_t$ is the expected incidence at time $t$ and $\beta(t, \tau)$ is the expected rate at which an individual infected $\tau$ time units ago infects individuals at calendar time $t$.
The model parameters can be used to define an effective reproduction number as either $\Re = \sum{\tau=0}^\infty \beta(t, \tau)$ or $\Re = \sum_{\tau=0}^\infty \beta(t+\tau, \tau)$~\autocite{gosticPractical}.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Renewal equations do not feature further in this thesis; for further details, including practical considerations when using them as the basis for inference, see \textcite{thompsonImproved,gosticPractical}.

@@

The infection process is hardly ever directly observed; however, we often observe a consequence of an infection.
\emph{Disease models} link the infection model to the number of these consequences, $W_t$ at time $t$.
For example, $W_t$ could be the incidence or prevalence of symptomatic individuals on day $t$.
$W_t$ will then be recorded through some system (\eg a testing programme) described by the observation model (see the next section).

At the core of many observation models is a \emph{convolution equation} or an approximation of it~\autocite[e.g.][]{brookmeyerBackcalculation,abbottEstimating,birrellBayesian}.
This equation describes the expectation of $W_t$, conditional on the incidence prior to time $t$.
In discrete time, it is:
\begin{align}
    \E(W_t \mid Z_1, Z_2, \dots) &= \zeta \sum_{i=0}^\infty Z_{t-i} f(i)
    \label{intro:eq:disease-model}
\end{align}
where $Z_t$ is the incidence at time $t$; $\zeta$ is the proportion of infections that result in $W_t$; and, conditional on being observed, $f(i)$ describes the (discretised) probability that an infection is included in the count $W_t$ at $i$ time units after the infection.
$f$ is an example of a \emph{delay function}, and $f$ is commonly the pdf of a distribution, or a transformation of it.

@@

In the application of the backcalculation to the CIS data, the delay function $f$ represents the duration of detectability.
Previous work has attempted to estimate $f$.

@@

\Textcite{mccabeCISincidence}, adapting the methodology \textcite{ealesAppropriately} developed for REACT, estimate $\Re$ and the growth rate from the CIS data.
They first estimate prevalence using summary statistics from the CIS, then derive $\Re$ and $r$ using a semi-mechanistic model fitted to their prevalence estimates.
% They then fit a semi-mechanistic model to the estimating prevalence 
% a posterior distribution of the parameters of a spline by fitting to the prevalence results via a binomial distribution.
% Their set-up penalizes the spline deviating from a constant growth rate.
% They calculate the growth rate from the derivative of the spline, an analytical transformation of the parameters. 
% Since, under reasonable model assumptions, there is a one-to-one correspondence between reproduction numbers and growth rates~\autocite{wallingaGI}, and they assume that the growth rate in incidence and prevalence is equal, this also estimates $\Re$.

@@

A secondary aim of this thesis is to improve our understanding of the epidemic in England over this period, and our understanding of the SARS-CoV-2 virus more broadly.
The results of the analyses I perform in pursuit of the primary aim will naturally contribute to this second aim.

@@

    \item In \cref{E-SEIR}, I develop a compartmental model for SARS-CoV-2 transmission, relying \emph{only} on CIS data. 
    I translate my estimate of the distribution of the duration of detectability into a progression through model states through an Erlang--Coxian distribution; this distribution is more flexible than standard approaches in the infectious disease literature.
    I include age-stratification in the transmission model, allowing estimation of age-specific transmission.
    I show the CIS data are sufficient for estimating incidence and transmission.
\end{enumerate}

@@

The secondary aim of this thesis is providing new estimates to improve our understanding of SARS-CoV-2 in general, and the epidemic in England over the period I study.

>>>>

\emph{Seroprevalence} surveys are more feasible.
Seroprevalence is the proportion of individuals in the population which have antibodies against a particular infection, indicating that the individual was previously infected~\autocite{cdcSeroprevalence}.
Many more individuals have ever been infected than are currently infected; therefore, seroprevalence surveys can be powered much more cheaply~\autocite{wuSeroprevSimulation}.
However, seroprevalence is a delayed measure because there is a lag between an individual being infected and them developing antibodies.

@@

Next, are semi-mechanistic models.
These relate incidence in the past to current (and future) incidence but do not explicitly represent the susceptible population.
A common type of semi-mechanistic models is that based on a \emph{renewal equation}~\autocite{bhattSemimechanistic,fraserRenewal}.
In discrete time, it takes the form:
\begin{align}
    \E(Z_t \mid Z_0, \dots, Z_{t-1}) &= \sum_{\tau=1}^t \beta(t,\tau) Z_{t-\tau} d \tau
    \label{intro:eq:renewal}
\end{align}
where $Z_t$ is the expected incidence at time $t$ and $\beta(t, \tau)$ is the expected rate at which an individual infected $\tau$ time units ago infects individuals at calendar time $t$.
The model parameters can be used to define an effective reproduction number as either $\Re = \sum_{\tau=0}^\infty \beta(t, \tau)$ or $\Re = \sum_{\tau=0}^\infty \beta(t+\tau, \tau)$~\autocite{gosticPractical}.
% A scaled measure of $I(t)$ (\eg hospital admissions or community test results) allows estimating $\beta$.
% The most common implementations are the R packages EpiNow2~\autocite{EpiNow2} and EpiEstim~\autocite{EpiEstim}.
Renewal equations do not feature further in this thesis; for further details, including practical considerations when using them as the basis for inference, see \textcite{thompsonImproved,gosticPractical}.

@@

The infection process is hardly ever directly observed; however, we often observe a consequence of an infection.
\emph{Disease models} link the infection model to the number of these consequences, $W_t$, at time $t$.
For example, $W_t$ could be the incidence or prevalence of symptomatic individuals on day $t$.
$W_t$ will then be recorded through some system (\eg a testing programme) described by the observation model (see the next section).

At the core of many observation models is a \emph{convolution equation} or an approximation of it~\autocite[e.g.][]{brookmeyerBackcalculation,abbottEstimating,birrellBayesian}.
This equation describes the expectation of $W_t$, conditional on the incidence prior to time $t$.
In discrete time, it is:
\begin{align}
    \E(W_t \mid Z_1, Z_2, \dots) &= \zeta \sum_{i=0}^\infty Z_{t-i} f(i)
    \label{intro:eq:disease-model}
\end{align}
where $Z_t$ is the incidence at time $t$; $\zeta$ is the proportion of infections that result in the consequence of interest; and, conditional on being observed, $f(i)$ describes the (discretised) probability that an infection is included in the count $W_t$ at $i$ time units after the infection.
$f$ is an example of a \emph{delay function}, and $f$ is commonly the pdf of a distribution, or a transformation of it.

@@

In the application of \cref{intro:eq:disease-model} to the CIS data, the delay function $f$ represents the duration of detectability.
Previous work has attempted to estimate $f$.

@@

\Textcite{mccabeCISincidence}, adapting the methodology \textcite{ealesAppropriately} developed for REACT, estimate $\Re$ and $r$ from the CIS data.
They first estimate prevalence using summary statistics from the CIS, then derive $\Re$ and $r$ using a semi-mechanistic model fitted to their prevalence estimates.
% They then fit a semi-mechanistic model to the estimating prevalence 
% a posterior distribution of the parameters of a spline by fitting to the prevalence results via a binomial distribution.
% Their set-up penalizes the spline deviating from a constant growth rate.
% They calculate the growth rate from the derivative of the spline, an analytical transformation of the parameters. 
% Since, under reasonable model assumptions, there is a one-to-one correspondence between reproduction numbers and growth rates~\autocite{wallingaGI}, and they assume that the growth rate in incidence and prevalence is equal, this also estimates $\Re$.

@@

The secondary aim of this thesis is to improve our understanding of the epidemic in England over this period, and our understanding of the SARS-CoV-2 virus more broadly.
The results of the analyses I perform in pursuit of the primary aim will naturally contribute to this second aim.

@@

    \item In \cref{E-SEIR}, I develop a compartmental model for SARS-CoV-2 transmission, relying \emph{only} on CIS data. 
    I translate my estimate of the distribution of the duration of detectability into a progression through model states.
    To do this, I use an Erlang--Coxian distribution; this distribution is more flexible than standard approaches in the infectious disease literature.
    I include age-stratification in the transmission model, allowing estimation of age-specific transmission.
    I show the CIS data are sufficient for estimating incidence and transmission.
\end{enumerate}

@@

The secondary aim of this thesis is providing new estimates to improve our understanding of the epidemic in England over the period I study.
